Type Flattening Obfuscation

### 2020-01-30 paper/tool.tex
update

Our effort is to attack the core of type reconstruction algorithms,
these are the data flow used to build type constraints, and the \emph{semantic gap} between types in the high-level language and their
representations in the low-level machine code.
Beside custom tricks which would be fully presented in another paper, \emph{uCc} uses extensively
obfuscation transformations based on Mixed Boolean Arithmetic expressions~\cite{eyrolles_obfuscation_2017,zhou_information_2007}.
To make some code diversity, the compiler is \emph{probabilistic}: given a source
code, \emph{uCc} generates each time a different (but computationally equivalent) output machine code.

We test outputs of \emph{uCc} against some decompilers (the test suite is available at
the repository). Table~\ref{description} describes briefly
tested functions and their types in source codes, table~\ref{flattening} shows results of using
$4$ decompilers to decompile binaries generated by \emph{uCc}.
No one between tested decompilers can recognize the precise signature of tested functions.
Basically, any argument is assigned by either \texttt{int64} or \texttt{uint64},
the highest types in the \emph{conversion rank} of integer types, we call this effect \emph{type flattening}.

We may note that if the tested functions are
compiled by, \emph{clang} for example, then their signatures can be recognized precisely
by the decompilers. Last but not least, \emph{uCc} makes no effort in making disassembling or function
boundary detection hard. Indeed, functions in tested binaries are perfectly detectable,
disassemblable and decompilable.

>>>>

Our effort is to attack the core of type reconstruction algorithms,
these are the data flow used to build type constraints, and the \emph{semantic gap} between types in the high-level language and their
representations in the low-level machine code.
Beside custom tricks, \emph{uCc} uses extensively
obfuscation transformations based on Mixed Boolean Arithmetic expressions~\cite{eyrolles_obfuscation_2017,zhou_information_2007}.
To make some code diversity, the compiler is \emph{probabilistic}: given a source
code, \emph{uCc} generates each time a different (but computationally equivalent) output machine code.

We test outputs of \emph{uCc} against some decompilers, table~\ref{description} describes briefly
tested functions and their types in source codes, table~\ref{flattening} shows results of
decompilers on binaries generated by \emph{uCc}. The effect of obfuscation is to make the types
of functions harder to be reconstructed. Indeed, the compilers know very little about the original
types. Basically, any argument in the function signature is simply assigned by either \texttt{int64}
or \texttt{uint64}, the highest types in the \emph{conversion rank} of integer types, we call this
effect \emph{type flattening}.

It may worth to note that the \emph{type flattening} is given by \emph{uCc}, not by the source code
complexity of tested functions. Indeed if they are compiled by, \emph{clang} or \emph{gcc} for example,
then their signatures are reconstructed precisely
by the decompilers. Last but not least, \emph{uCc} makes no effort in making disassembling or function
boundary detection hard: functions in tested binaries are perfectly detectable,
disassemblable and decompilable.
